### 2024-09-26 manuscript.tex
Improve style

% Ramadoss et al. (2020), Simulating mixed-phase cloud properties with ICON around the CAPRICORN ﬁeld campaign at the kilometre scale [poster]

@@

\documentclass[12pt,a4paper]{article}
\usepackage{fontspec}
\usepackage[margin=2cm]{geometry}
\usepackage[citecolor=black]{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{doi}
\usepackage{titlesec}
\usepackage{caption}
\DeclareCaptionLabelSeparator{pipe}{ | }
\captionsetup{labelfont={bf,sf},labelsep=pipe}
\usepackage{graphicx}
\setsansfont{Source Sans Pro}
\setmainfont{Source Serif Pro}
\hypersetup{
	colorlinks=true,
	urlcolor=blue,
	pdftitle={Ship and ground-based lidar and radiosonde evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON and the ERA5 and MERRA-2 reanalyses},
	pdfauthor={Peter Kuma, Frida A.-M. Bender, Adrian J. McDonald and Thorsten Mauritsen}
}
\bibliographystyle{abbrvnat}
\titleformat*{\section}{\large\bfseries\sffamily}
\titleformat*{\subsection}{\normalsize\bfseries\sffamily}
\titlespacing*{\section}{0pt}{12pt}{6pt}
\titlespacing*{\subsection}{0pt}{8pt}{4pt}
%\renewcommand{\bibsection}{\section*{Publications}}
\parskip=6pt
\parindent=0pt

@@

\begin{center} \Large \sffamily\textbf{Ship and ground-based lidar and radiosonde evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON and the ERA5 and MERRA-2 reanalyses}\\[0.4cm]
\large Peter Kuma$^\mathsf{1,}$*, Frida A.-M. Bender$^\mathsf{1}$, Adrian J. McDonald$^\mathsf{2}$ and Thorsten Mauritsen$^\mathsf{1}$\\[0.4cm]
\small
$^\mathsf{1}$Department of Meteorology (MISU) and Bolin Centre for Climate Research, Stockholm University, Stockholm SE-106 91, Sweden\\[0.2cm]
$^\mathsf{2}$School of Physical and Chemical Sciences, University of Canterbury, Christchurch, New Zealand\\[0.2cm]
*Corresponding author: Peter Kuma (\href{mailto:peter.kuma@misu.su.se}{peter.kuma@misu.su.se})\\[0.4cm]
\large \today\\[0.4cm]

@@

Previous studies have identified substantial large-scale biases in model clouds
over the Southern Ocean (SO), affecting sea surface temperature and the Earth's
albedo overall. Our aim is to quantify how well the high-resolution ICON model
is simulating clouds in this region, particularly in light of the fact that
subgrid-scale clouds are not parametrised in this model. This region is mostly
dominated by boundary layer clouds generated by shallow convection, and these
are problematic to observe by spaceborne lidar and radars, which are affected
by attenuation by overlapping and thick clouds and ground clutter,
respectively. Therefore, we choose to use a large set of ship-based
observations conducted with ceilometers and lidars on board of the Research
Vessel (RV) \textit{Polarstern} and other voyages. Altogether, we analyse over
2000 days (about 6 years) of data from 31 voyages and 1 sub-antarctic station
covering diverse longitudes of the SO. To achieve a like-for-like comparison
with the model, we use a ground-based lidar simulator called the Automatic
Lidar and Ceilometer Framework (ALCF). We contrast the results with the ECMWF
Reanalysis 5 (ERA5) and the Modern-Era Retrospective analysis for Research and
Applications, Version 2 (MERRA-2).

@@

The main aim of our research is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst (DWD) and the Max-Planck-Institute
for Meteorology (MPI-M). Previous studies have identified substantial
large-scale biases in climate model clouds over the Southern Ocean (SO),
affecting sea surface temperature and the Earth’s albedo overall. Our aim is to
quantify how well the high-resolution ICON model is simulating clouds in this
region, particularly in light of the fact that subgrid-scale clouds are not
parameterised in this model. This region is mostly dominated by boundary layer
clouds generated by shallow convection, and these are problematic to observe by
spaceborne lidar and radars, which are affected by attenuation by overlapping
and thick clouds and ground clutter, respectively.  Specifically, the radar on
CloudSat and lidar on CALIPSO are affected by the abovementioned issues,
resulting in a strong underestimation of cloud occurrence below 2 km relative
to ground-based lidar observations \citep{mcerlich2021}.  This, in turn, can
lead to systematic biases in low clouds in climate models, which are frequently
evaluated against CloudSat--CALIPSO products. Reanalyses can also suffer from
cloud biases, as these are usually parametrised in their atmospheric component,
and also in regions where input observations are sparse.  This makes them a
problematic reference for clouds over the SO, and any biases relative to a
reanalysis should be interpreted with caution. Instead, we chose to use a large
set of ship-based observations conducted with ceilometers and lidars on board
of the RV \emph{Polarstern} and other voyages and stations as a reference for
the model evaluation.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/map_fig.pdf}
\caption{
\textbf{(a)} A map showing the tracks of 31 voyages of RV \emph{Polarstern},
RSV \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B. Palmer},
and HMNZS \emph{Wellington} and one sub-Antarctic station (Macquarie Island)
analysed here. The tracks cover Antarctic sectors south of South America, the
Atlantic Ocean, Africa, Australia, and New Zealand in the years 2010--2021
(inclusive).  The dotted and dashed lines at 40°S and 55°S delineate the
Southern Ocean area of our analysis and its partitioning into two subsets,
respectively.  A photo of \textbf{(b)} RV \emph{Polarstern} (© Folke Mehrtens,
Alfred-Wegener-Institut), \textbf{(c)} Lufft CHM 15k installed on RV
\emph{Tangaroa} (© Peter Kuma, University of Canterbury), \textbf{(d)} Vaisala
CL51 installation at the Macquarie Island station (© Jeﬀ Aquilina, Bureau of
Meteorology).
% TODO: Ask for photo permission and info about license.
}
\label{fig:map}
\end{figure}

@@

On the RV \emph{Polarstern} voyages, regular human-performed synoptic
observations were available and included precipitation presence or absence and
type. We used this dataset to train a convolutional artiﬁcial neural network
(ANN) of the U-Net type \citep{ronneberger2015} to recognise profiles with
precipitation from lidar backscatter (Fig. \ref{fig:ann}a), implemented in the
TensorFlow ANN framework \citep{tensorflow}. Samples of short time intervals
(10 min) of near-surface lidar backscatter (0–250 m) were classified as clear,
rain, snow, and fog, using the synoptic observations as a training dataset
(Fig.  \ref{fig:ann}b). From these, a binary, mutually exclusive classification
of profiles as precipitating (rain or snow) or dry (clear or fog) was derived.
For detecting model and reanalysis precipitation, we used a fixed threshold for
surface precipitation flux of 0.1 mm h$^{-1}$ (the ANN was not used).

The ANN achieved 65\% sensitivity and 87\% speciﬁcity when the true positive
rate (26\%) was made to match observations. The receiver operating
characteristic curve is shown in Fig. \ref{fig:ann}c. We considered these rates
satisfactory for the purpose of filtering precipitation profiles. Fig.
\ref{fig:ann}d shows examples of the predicted precipitation compared to
human-performed observations.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/cyc_stab_dist.pdf}
\caption{
Geographical distribution of \textbf{(a, b)} cyclonic days and \textbf{(b, d)}
relatively stable (LTS > 12 K) time steps in \textbf{(a, c)} ERA5 in years
2010--2013 (inclusive) and \textbf{(b, d)} ICON in model years 2021--2023 (free
running). Cyclonic days are expressed as a fraction of the number of days with
cyclonic activity, defined as grid points located within a double radius of any
cyclone on a given day (UTC), as identified by CyTRACK.
}
\label{fig:cyclone-stability}
\end{figure}

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, and two reanalyses, MERRA-2 and ERA5 (Fig.
\ref{fig:cloud-occurrence-panel}). We aggregated the sources by calculating the
averages and percentiles of all individual profiles (Fig.
\ref{fig:cloud-occurrence}). The analysis shows that the total cloud fraction
(determined as the fraction of proﬁles with clouds at any height in the lidar
cloud mask) is underestimated in ICON and reanalyses by about 10\% and 20\%,
respectively.

@@

The ICON comparison is limited by the fact that the model is free-running.
Thus, the vertical proﬁles are not expected to represent the same weather
conditions as in observations, but long-term statistical comparison is still
possible. Only proﬁles with the same sea ice conditions (present or not) were
included to avoid comparing sea ice with open sea conditions.

@@

We analysed the daily cloud cover (total cloud fraction) distribution. This is
a measure of cloudiness, irrespective of height. A cloud detected at any height
means that the lidar profile was classified as cloudy; otherwise, it was
classified as a clear sky. When all profiles in a day are aggregated together,
the cloud cover for the day is defined as the fraction of cloudy profiles in
the total number of profiles, expressed in octas.

In Fig.  \ref{fig:cloud-cover} we show the results for the same subsets as
analysed above. Observations show the greatest representation of high cloud
cover (5--8 octas), peaking at 7 octas. This is not represented by ICON or the
reanalyses.  While ICON is the closest, it tends to be 1 octa clearer the
observations, peaking at 6 octas, and highly understimating days with 8 octas.
Overall, the reanalyses show results similar to each other, underestimating
cloud cover by about 2 octas, and strongly understimating days with 7 and 8
octas. Of the two reanalyses, MERRA-2 shows slightly higher cloud cover and
thus more consistent with observations.

When analysed by subsets, the cyclonic subset shows the highest cloud cover,
with 8 octas occurring on one half of such days (Fig. \ref{fig:cloud-cover}d).
This is not represented by ICON or the reanalyses at all. Interestingly, clear
sky days (0 octas) also have a local maximum peaking at about 15\% in this
subset.  When we contrast the low and high latitude zones, we see that the high
latitude zone tends to have greater cloud cover, peaking at 8 octas (Fig.
\ref{cloud-cover}c). The high latitude also has almost no clear sky or small
cloud cover cases (0--4 octas). ICON and the reanalyses represent at least this
characteristic of the distribution well for 0--3 octas, but otherwise show
bises similar to the general case. One of the greatest biases is present in
ERA5 in the unstable subset, where it peaks at 3 octas, whereas the
observations peak at 7 octas and show negligible cloud cover below 5 octas.

@@

\subsection{Thermodynamic proﬁles}

We analysed about 2300 radiosonde proﬁles south of 40°S from the 24 RV
\emph{Polarstern} voyages, MARCUS, NBP1704, TAN1702, and TAN1802. Spatially and
temporally colocated profiles were taken from ICON and the reanalyses. Because
the time period of the ICON model output was different from the observations,
model time was chosen to be the same as the radiosonde launch time relative to
the start of the year. The profiles were partitioned into the same subsets as
the lidar observations: all, 40--55°S, 55+°S, cyclonic, non-cyclonic, stable,
and unstable. The groups 40--55°S and 55+°S, cyclonic and non-cyclonic, and
stable and unstable were mutually exclusive in pairs. Apart from relative
humidity, we focus on comparing virtual potential temperature ($\theta_v$) due
to its role in low-level tropospheric stability, being one of the primary
factors affecting shallow convection and the associated low-level cloud
formation and dissipation. The observed and model profiles of virtual potential
temperature are shown in Fig. \ref{fig:potential-temperature}.

@@

We analysed a total of about 2421 days of lidar and radiosonde data from 31
voyages and campaigns and one subantarctic station data, covering the Atlantic,
Australian and New Zealand sectors of the SO over the span of 10 years. This
dataset, together with the use of the ALCF lidar simulator, provided a fairly
comprehensive basis for evaluating SO cloud and thermodynamic profile biases in
the GSRM ICON and the ERA5 and MERRA-2 reanalyses. This provides a unique
evaluation perspective different from satellite observations, one which is
more suitable for evaluating bounary layer clouds, which are predominant in
this region. Furthermore, we subsetted our dataset by low and high latitude
bands, cyclonic activity and stability, in order to identify how these
conditions relate the the biases. Our main finding corroborates previous
findings of large boundary layer cloud biases in models, and their subsequent
effect on the radiative transfer. This also applies to the new ICON GSRM,
but the biases are lower than in the reanalyses, despite the reanalyses having
the advantage of assimilation of the observed meteorological conditions.
The GSRM has, on the other hand, the advange of a much higher spatial
resolution and, to some extent, explicit calculation of traditionally
subgrid-scale processes such as convection.

@@

We compared radiosonde proﬁles acquired on the RV \emph{Polarstern} voyages
with ICON.  The ICON model exhibits smaller internal variability than
observations, and its LCL level tends to be higher, which might explain why a
peak of cloud occurrence in observations is located at the surface, while in
ICON it is located higher, corresponding to the LCL.

@@

\section*{Acknowledgements}

We acknowledge the RV Polarstern datasets provided by the Alfred Wegener
Institute, RSV Aurora Australis datasets provided by the Australian Antarctic
Program and the University of Canterbury (UC), RV Tangaroa datasets provided by
the National Institute of Water and Atmospheric Research and the UC, RV
Nathaniel B. Palmer dataset provided by the National Science Foundation,
Cooperative Institute for Research in Environmental Sciences, University of
Colorado and the UC, HMNZS Wellington dataset provided by the Royal New Zealand
Navy and the UC and a Macquarie Island dataset provided by the Atmospheric
Radiation Measurement and the UC. We thank the scientiﬁc staff, the crew, and
everyone involved in collecting data on the voyages and Macquarie Island,
especially Kelly Schick, Peter Guest, John J. Cassano, Simon Parsons, Graeme
Plank, Sally Garrett, Jamie Halla, Sean Hartery, Mike J.  Harvey, Simon P.
Alexander and Alex Schuddeboom. We acknowledge the reanalysis dataset ERA5
provided by the Copernicus Climate Change Service and MERRA-2 provided by the
Global Modeling and Assimilation Office. We also acknowledge the AMSR Uniﬁed
and SMMR and SSM/I-SSMIS datasets provided by the National Snow and Ice Data
Center.

@@

\footnotesize
\setlength{\bibsep}{0.0pt}
\bibliography{manuscript}

>>>>

% Ramadoss et al. (2020), Simulating mixed-phase cloud properties with ICON around the CAPRICORN field campaign at the kilometre scale [poster]

@@

\documentclass[12pt,a4paper]{article}
\usepackage{fontspec}
\usepackage[margin=2cm]{geometry}
\usepackage[citecolor=black]{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{doi}
\usepackage{titlesec}
\usepackage{caption}
\DeclareCaptionLabelSeparator{pipe}{ | }
\captionsetup{labelfont={bf,sf},labelsep=pipe}
\usepackage{graphicx}
\usepackage{lineno}
\linenumbers
\setsansfont{Source Sans Pro}
\setmainfont{Crimson Text}
\setmonofont[Scale=0.8]{Space Mono}
\hypersetup{
	colorlinks=true,
	urlcolor=blue,
	pdftitle={Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses},
	pdfauthor={Peter Kuma, Frida A.-M. Bender, Adrian J. McDonald and Thorsten Mauritsen}
}
\bibliographystyle{abbrvnat}
\titleformat*{\section}{\large\bfseries\sffamily}
\titleformat*{\subsection}{\normalsize\bfseries\sffamily}
\titlespacing*{\section}{0pt}{12pt}{6pt}
\titlespacing*{\subsection}{0pt}{10pt}{4pt}
%\renewcommand{\bibsection}{\section*{Publications}}
\parskip=6pt
\parindent=0pt

@@

\fontsize{13pt}{15pt}\selectfont

\begin{center} \Large \sffamily\textbf{Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON and the ERA5 and MERRA-2 reanalyses}\\[0.4cm]
\large Peter Kuma$^\mathsf{1,}$*, Frida A.-M. Bender$^\mathsf{1}$, Adrian J. McDonald$^\mathsf{2}$ and Thorsten Mauritsen$^\mathsf{1}$\\[0.4cm]
\small
$^\mathsf{1}$Department of Meteorology (MISU) and Bolin Centre for Climate Research, Stockholm University, Stockholm SE-106 91, Sweden\\[0.2cm]
$^\mathsf{2}$School of Physical and Chemical Sciences, University of Canterbury, Christchurch, New Zealand\\[0.2cm]
*Corresponding author: Peter Kuma (\href{mailto:peter.kuma@misu.su.se}{peter.kuma@misu.su.se})\\[0.4cm]
\large \today\\[0.4cm]

@@

Previous studies have identified substantial large-scale biases in model clouds
over the Southern Ocean (SO), affecting sea surface temperature and the Earth's
albedo overall. Our aim is to quantify how well the high-resolution ICON model
is simulating clouds in this region, particularly in light of the fact that
subgrid-scale clouds are not parametrised in this model. This region is mostly
dominated by boundary layer (BL) clouds generated by shallow convection, and
these are problematic to observe by spaceborne lidar and radars, which are
affected by attenuation by overlapping and thick clouds and ground clutter,
respectively. Therefore, we choose to use a large set of ship-based
observations conducted with ceilometers and lidars on board of the Research
Vessel (RV) \textit{Polarstern} and other voyages. Altogether, we analyse over
2000 days (about 6 years) of data from 31 voyages and 1 sub-antarctic station
covering diverse longitudes of the SO. To achieve a like-for-like comparison
with the model, we use a ground-based lidar simulator called the Automatic
Lidar and Ceilometer Framework (ALCF). We contrast the results with the ECMWF
Reanalysis 5 (ERA5) and the Modern-Era Retrospective analysis for Research and
Applications, Version 2 (MERRA-2).

@@

The main aim of our research is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst (DWD) and the Max-Planck-Institute
for Meteorology (MPI-M). Previous studies have identified substantial
large-scale biases in climate model clouds over the Southern Ocean (SO),
affecting sea surface temperature and the Earth’s albedo overall. Our aim is to
quantify how well the high-resolution ICON model is simulating clouds in this
region, particularly in light of the fact that subgrid-scale clouds are not
parameterised in this model. This region is mostly dominated by BL clouds
generated by shallow convection, and these are problematic to observe by
spaceborne lidar and radars, which are affected by attenuation by overlapping
and thick clouds and ground clutter, respectively.  Specifically, the radar on
CloudSat and lidar on CALIPSO are affected by the abovementioned issues,
resulting in a strong underestimation of cloud occurrence below 2 km relative
to ground-based lidar observations \citep{mcerlich2021}.  This, in turn, can
lead to systematic biases in low clouds in climate models, which are frequently
evaluated against CloudSat--CALIPSO products. Reanalyses can also suffer from
cloud biases, as these are usually parametrised in their atmospheric component,
and also in regions where input observations are sparse.  This makes them a
problematic reference for clouds over the SO, and any biases relative to a
reanalysis should be interpreted with caution. Instead, we chose to use a large
set of ship-based observations conducted with ceilometers and lidars on board
of the RV \emph{Polarstern} and other voyages and stations as a reference for
the model evaluation.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/map_fig.pdf}
\caption{
\textbf{(a)} A map showing the tracks of 31 voyages of RV \emph{Polarstern},
RSV \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B. Palmer},
and HMNZS \emph{Wellington} and one sub-Antarctic station (Macquarie Island)
analysed here. The tracks cover Antarctic sectors south of South America, the
Atlantic Ocean, Africa, Australia, and New Zealand in the years 2010--2021
(inclusive).  The dotted and dashed lines at 40°S and 55°S delineate the
Southern Ocean area of our analysis and its partitioning into two subsets,
respectively.  A photo of \textbf{(b)} RV \emph{Polarstern} (© Folke Mehrtens,
Alfred-Wegener-Institut), \textbf{(c)} Lufft CHM 15k installed on RV
\emph{Tangaroa} (© Peter Kuma, University of Canterbury), \textbf{(d)} Vaisala
CL51 installation at the Macquarie Island station (© Jeff Aquilina, Bureau of
Meteorology).
% TODO: Ask for photo permission and info about license.
}
\label{fig:map}
\end{figure}

@@

On the RV \emph{Polarstern} voyages, regular human-performed synoptic
observations were available and included precipitation presence or absence and
type. We used this dataset to train a convolutional artificial neural network
(ANN) of the U-Net type \citep{ronneberger2015} to recognise profiles with
precipitation from lidar backscatter (Fig. \ref{fig:ann}a), implemented in the
TensorFlow ANN framework \citep{tensorflow}. Samples of short time intervals
(10 min) of near-surface lidar backscatter (0–250 m) were classified as clear,
rain, snow, and fog, using the synoptic observations as a training dataset
(Fig.  \ref{fig:ann}b). From these, a binary, mutually exclusive classification
of profiles as precipitating (rain or snow) or dry (clear or fog) was derived.
For detecting model and reanalysis precipitation, we used a fixed threshold for
surface precipitation flux of 0.1 mm h$^{-1}$ (the ANN was not used).

The ANN achieved 65\% sensitivity and 87\% specificity when the true positive
rate (26\%) was made to match observations. The receiver operating
characteristic curve is shown in Fig. \ref{fig:ann}c. We considered these rates
satisfactory for the purpose of filtering precipitation profiles. Fig.
\ref{fig:ann}d shows examples of the predicted precipitation compared to
human-performed observations.

@@

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/cyc_stab_dist.pdf}
\caption{
Geographical distribution of \textbf{(a, b)} cyclonic days and \textbf{(b, d)}
relatively stable (LTS > 12 K) time steps in \textbf{(a, c)} ERA5 in years
2010--2013 (inclusive) and \textbf{(b, d)} ICON in model years 2021--2023 (free
running). Cyclonic days are expressed as a fraction of the number of days with
cyclonic activity, defined as grid points located within a double radius of any
cyclone on a given day (UTC), as identified by CyTRACK.
}
\label{fig:cyclone-stability}
\end{figure}

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, and two reanalyses, MERRA-2 and ERA5 (Fig.
\ref{fig:cloud-occurrence-panel}). We aggregated the sources by calculating the
averages and percentiles of all individual profiles (Fig.
\ref{fig:cloud-occurrence}). The analysis shows that the total cloud fraction
(determined as the fraction of profiles with clouds at any height in the lidar
cloud mask) is underestimated in ICON and reanalyses by about 10\% and 20\%,
respectively.

@@

The ICON comparison is limited by the fact that the model is free-running.
Thus, the vertical profiles are not expected to represent the same weather
conditions as in observations, but long-term statistical comparison is still
possible. Only profiles with the same sea ice conditions (present or not) were
included to avoid comparing sea ice with open sea conditions.

@@

We analysed the daily cloud cover (total cloud fraction) distribution. This is
a measure of cloudiness, irrespective of height. A cloud detected at any height
means that the lidar profile was classified as cloudy; otherwise, it was
classified as a clear sky. When all profiles in a day are aggregated together,
the cloud cover for the day is defined as the fraction of cloudy profiles in
the total number of profiles, expressed in oktas.

In Fig.  \ref{fig:cloud-cover} we show the results for the same subsets as
analysed above. Observations show the greatest representation of high cloud
cover (5--8 oktas), peaking at 7 oktas. This is not represented by ICON or the
reanalyses.  While ICON is the closest, it tends to be 1 okta clearer the
observations, peaking at 6 oktas, and highly understimating days with 8 oktas.
Overall, the reanalyses show results similar to each other, underestimating
cloud cover by about 2 oktas, and strongly understimating days with 7 and 8
oktas. Of the two reanalyses, MERRA-2 shows slightly higher cloud cover and
thus more consistent with observations.

When analysed by subsets, the cyclonic subset shows the highest cloud cover,
with 8 oktas occurring on one half of such days (Fig. \ref{fig:cloud-cover}d).
This is not represented by ICON or the reanalyses at all. Interestingly, clear
sky days (0 oktas) also have a local maximum peaking at about 15\% in this
subset.  When we contrast the low and high latitude zones, we see that the high
latitude zone tends to have greater cloud cover, peaking at 8 oktas (Fig.
\ref{fig:cloud-cover}c). The high latitude also has almost no clear sky or
small cloud cover cases (0--4 oktas). ICON and the reanalyses represent at
least this characteristic of the distribution well for 0--3 oktas, but
otherwise show bises similar to the general case. One of the greatest biases is
present in ERA5 in the unstable subset, where it peaks at 3 oktas, whereas the
observations peak at 7 oktas and show negligible cloud cover below 5 oktas.

@@

\subsection{Thermodynamic profiles}

We analysed about 2300 radiosonde profiles south of 40°S from the 24 RV
\emph{Polarstern} voyages, MARCUS, NBP1704, TAN1702, and TAN1802. Spatially and
temporally colocated profiles were taken from ICON and the reanalyses. Because
the time period of the ICON model output was different from the observations,
model time was chosen to be the same as the radiosonde launch time relative to
the start of the year. The profiles were partitioned into the same subsets as
the lidar observations: all, 40--55°S, 55+°S, cyclonic, non-cyclonic, stable,
and unstable. The groups 40--55°S and 55+°S, cyclonic and non-cyclonic, and
stable and unstable were mutually exclusive in pairs. Apart from relative
humidity, we focus on comparing virtual potential temperature ($\theta_v$) due
to its role in low-level tropospheric stability, being one of the primary
factors affecting shallow convection and the associated low-level cloud
formation and dissipation. The observed and model profiles of virtual potential
temperature are shown in Fig. \ref{fig:potential-temperature}.

@@

We analysed a total of about 2421 days of lidar and radiosonde data from 31
voyages and campaigns and one subantarctic station data, covering the Atlantic,
Australian and New Zealand sectors of the SO over the span of 10 years. This
dataset, together with the use of the ALCF lidar simulator, provided a fairly
comprehensive basis for evaluating SO cloud and thermodynamic profile biases in
the GSRM ICON and the ERA5 and MERRA-2 reanalyses. This provides a unique
evaluation perspective different from satellite observations, one which is more
suitable for evaluating bounary layer clouds, which are predominant in this
region. Furthermore, we subsetted our dataset by low and high latitude bands,
cyclonic activity and stability, in order to identify how these conditions
relate the the biases. Our main finding corroborates previous findings of large
BL cloud biases in models, and their subsequent effect on the radiative
transfer. This also applies to the new ICON GSRM, but the biases are lower than
in the reanalyses, despite the reanalyses having the advantage of assimilation
of the observed meteorological conditions.  The GSRM has, on the other hand,
the advange of a much higher spatial resolution and, to some extent, explicit
calculation of traditionally subgrid-scale processes such as convection.

@@

We compared radiosonde profiles acquired on the RV \emph{Polarstern} voyages
with ICON.  The ICON model exhibits smaller internal variability than
observations, and its LCL level tends to be higher, which might explain why a
peak of cloud occurrence in observations is located at the surface, while in
ICON it is located higher, corresponding to the LCL.

@@

\fontsize{12pt}{14pt}\selectfont
\section*{Acknowledgements}

We acknowledge the RV Polarstern datasets provided by the Alfred Wegener
Institute, RSV Aurora Australis datasets provided by the Australian Antarctic
Program and the University of Canterbury (UC), RV Tangaroa datasets provided by
the National Institute of Water and Atmospheric Research and the UC, RV
Nathaniel B. Palmer dataset provided by the National Science Foundation,
Cooperative Institute for Research in Environmental Sciences, University of
Colorado and the UC, HMNZS Wellington dataset provided by the Royal New Zealand
Navy and the UC and a Macquarie Island dataset provided by the Atmospheric
Radiation Measurement and the UC. We thank the scientific staff, the crew, and
everyone involved in collecting data on the voyages and Macquarie Island,
especially Kelly Schick, Peter Guest, John J. Cassano, Simon Parsons, Graeme
Plank, Sally Garrett, Jamie Halla, Sean Hartery, Mike J.  Harvey, Simon P.
Alexander and Alex Schuddeboom. We acknowledge the reanalysis dataset ERA5
provided by the Copernicus Climate Change Service and MERRA-2 provided by the
Global Modeling and Assimilation Office. We also acknowledge the AMSR Unified
and SMMR and SSM/I-SSMIS datasets provided by the National Snow and Ice Data
Center.

@@

%\footnotesize
\fontsize{11pt}{13pt}\selectfont
\setlength{\bibsep}{0.0pt}
\bibliography{manuscript}
